Hierarchical recurrent models reveal latent states of neural activity in \textit{C. elegans}

### 2017-12-21 drafts/elife/draft.tex
recurrent results section

@@

>>>>

The final and most critical component of the dynamical model for
neural activity in the head ganglia is the discrete transition
distribution.  In contrast to standard SLDS, we allow these transition
probabilities to depend on the location in continuous state space (the
\emph{recurrent} dependency) as well as the preceding discrete state
(the \emph{Markovian} dependency).  Fig.~\ref{fig:recurrent}a shows
the contribution of the Markovian term.  Each cell in this matrix
denotes the number of transitions from one state to another. We are
only counting transitions in which the state changes, hence the
diagonal is zero.  We have permuted the states to emphasize the
largely feed-forward structure.

Fig.~\ref{fig:recurrent}b shows the strongest entries in this matrix
in an alternative form.  We have illustrated each state as a node in a
graph, and labeled the states based on their correspondence with the
manually labeled states of~\citet{kato2015global} (see
Fig.~\ref{fig:syllables}d).  This reveals a chain-like transition
structure in which slow and forward (\textsf{FWD}) crawling are linked
to reverse (\textsf{REV1}, \textsf{REV2}) and sustained reverse
(\textsf{REVSUS}) via dorsal turns (\textsf{DT}). From sustained
reverse crawling, the worm may execute either dorsal or ventral turns
(\textsf{VT}).

To accurately model these dynamics, the changes must occur at precise
locations in continuous latent space.  Namely,
Fig.~\ref{fig:syllables}a shows that the loops tend to start and end
at the origin, and we rarely see transitions when the continuous
latent state is far from zero. This manifests in signature duration
distributions for the various states.  Fig.~\ref{fig:recurrent}c plots
these distributions for each of the eight states.  Superimposed on
top, we show the predicted distributions with the recurrent SLDS
(solid black lines) compared with the predictions of the standard,
Markovian SLDS (dotted black lines). The Markovian model knows nothing
of the continuous latent state and implies a geometric duration
distribution with monotonically decreasing probability. By contrast,
the recurrent model learns to transition near the origin, which
gives rise to characteristic duration distributions that closely
match the empirical distributions.

@@

Having demonstrated the utility of
